### 2011-04-03 docs/enthalpy/enthalpy.tex
Added more stuff about derivation of constraint.

\section{Enthalpy Evolution}

In the low Mach number formulation, the continuity and momentum
equations together with the constraint on the velocity field
completely determine the solution.
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + 
\rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - 
  \frac{1}{\rho}\nabla\pi - 
  \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}
\end{eqnarray}

The addition of an energy equation over-specifies the problem, but its
solution can help us in providing a diagnostic capability for the
solution.  Additionally, work from combustion suggests that low Mach
number methods perform better when considering reactions if we use
an enthalpy equation to determine the temperature.

The velocity constraint equation was constructed by diffrentiating 
the pressure along particle paths, using the equation of state.

The full enthalpy equation, with no approximations, appears as:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp}{Dt} + \rho\Hnuc \label{eq:enthalpy}
\end{equation}
Here, $h = e + p/\rho$ is the specific enthalpy, with $e$ the specific
internal energy.  In the low Mach number formulation, we replace $p$
with $p_0$ in the $Dp/Dt$ term, however, the definition of enthalpy
implicitly contains a pressure.  When calling the equation of state,
we take $h$ and $\rho$ as inputs.  The equation of state is expressed
in terms of $T$ and $\rho$, so it iterates until it finds the $h$ that
we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/\rho$,
where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
$p_\mathrm{EOS}$ may not be equal to $p_0$---this causes us to drive
off of the constraint. \MarginPar{not completely sure what to say here}

@@

\item We always say that to substituting $p$ by $p_0$ in the enthalpy equation
   is what filters the acoustics from the system---did we ever prove this?

>>>>

\section{Evolution Equations}

The compressible and low Mach number formulations of the governing
equations both share the unapproximated continuity and momentum equations
shown here (where $p = p_0 + \pi$).
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + 
\rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - 
  \frac{1}{\rho}\nabla\pi - 
  \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}
\end{eqnarray}

In the compressible formulation we complete the system with an energy
equation as well as an equation of state; in the low Mach number
formulation we can derive a constraint on the velocity by 
setting $p_{EOS} = p(\rho,h,X_k) = p_0$ and
differentiating the equation of state along particle paths.
In this case adding the energy equation would over-constrain the system, 
but its solution can help us in providing a diagnostic capability for the
solution.  We can write the energy equation in terms of internal energy, $e$,
or enthalpy, $h$; these two equations are analytically equivalent.
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} + \nabla\cdot(\rho h\Ub) 
 &=& \frac{Dp}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \nabla\cdot(\rho e \Ub)
 &=& - p\nabla\cdot \Ub + \rho\Hnuc \nonumber 
\end{eqnarray}
In low Mach number combustion, $Dp/Dt = 0,$ which makes the enthalpy equation
preferable to the internal energy equation because we don't need to
evaluate $p \nabla \cdot \Ub.$

\section{Derivation of Constraint Using Enthalpy Equation}

The velocity constraint equation was constructed by differentiating 
the pressure along particle paths, using the equation of state.  
Specifically, we wrote $p_{EOS} = p(\rho,T,X_k) = p_0$
and expressed $DT/Dt$ in terms of $Dh/Dt:$
\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
- \sum_k \rho \xi_k \omegadot_k + \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_p = \left.\partial h/\partial T\right|_{p,X_k}$ is the
specific heat at constant pressure,
$\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
and $h_p = \left.\partial h/\partial p\right|_{T,X_k}$
for convenience.

Then we derived
\begin{eqnarray*}
\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
- \frac{D p}{D t} + \frac{p_T}{\rho c_p}
  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho H \right)
+ \sum_k p_{X_k} \omegadot_k \right)  \\
                 &=& \frac{1}{\rho p_\rho}
  \left( \frac{p_T}{\rho c_p}(1  - \rho h_p) - 1 \right) \frac{D p}{D t}
 + \frac{1}{\rho p_\rho} \left(
  \frac{p_T}{\rho c_p} (\rho H - \rho \sum_k \xi_k   \omegadot_k)
                               + \sum_k p_{X_k} \omegadot_k \right) \enskip .
\end{eqnarray*}
When we derived this expression we explicitly retained the dependence of $h$ on $p$,
as shown by the presence of the $h_p$ term.

Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
\nabla p_0$, and, recalling the definition of $H$, the divergence constraint can
be written
\begin{equation}
\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
\frac{1}{\rho p_\rho} \left(
   \frac{p_T}{\rho c_p} \left(\nabla \cdot (\kappa \nabla T)
  - \sum_k\rho (q_k + \xi_k) \omegadot_k \right)
 + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
\end{equation}
where we define
\begin{equation}
\alpha(\rho,T) \equiv - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \enskip . \label{eq:alphadef}
\end{equation}

\section{Derivation of Constraint Using Energy Equation}

We could also derive the constraint by again differentiating $p_{EOS} = p(\rho,T,X_k) = p_0$
but this time writing $DT/Dt$ in terms of $De/Dt:$

\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_v} \left( -e_p \frac{D p}{D t}
- \sum_k e_k \frac{D X_k}{D t} - p \nabla \cdot U 
+ \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
specific heat at constant volume,
and $e_p = \left.\partial e/\partial p\right|_{T,X_k}$
for convenience.

Then...

\section{Enthalpy vs Energy Equation}

The full enthalpy equation, with no approximations, appears as:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp}{Dt} + \rho\Hnuc \label{eq:enthalpy}
\end{equation}
Here, $h = e + p/\rho$ is the specific enthalpy, with $e$ the specific
internal energy.  In the low Mach number formulation, we replace $p$
with $p_0$ in the $Dp/Dt$ term, however, the definition of enthalpy
implicitly contains a pressure.  When calling the equation of state,
we take $h$ and $\rho$ as inputs.  The equation of state is expressed
in terms of $T$ and $\rho$, so it iterates until it finds the $h$ that
we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/\rho$,
where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
$p_\mathrm{EOS}$ may not be equal to $p_0$---this may be what
causes us to drive off of the constraint.

@@

\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
constraint, 2) in the enthalpy evolution equation.   Are these substitutions
consistent with each other?  If they are not consistent, which of these 
substitutions is responsible for filtering the acoustics from the system?
